---
title: "Homework 3"
output:
  html_document: default
  pdf_document: default
date: 'Deadline: 9 February, 12:00'
subtitle: 'Linguistic Data: Quantitative Analysis and Visualisation'
---

The solutions should be submitted via GitHub.  

## Part 1. A preliminary training  

*Do not use R (RStudio) to solve problems in Part 1. Answers won't be evaluated. *

### Problem 1

Look at the following histogram and answer the questions.

```{r, echo=FALSE}
set.seed(1234)
x <- rnorm(20, mean=3, sd=6)
hist(x, col="tomato", main="")
```

a. What is the proportion of values in the sample that exceed $5$? Explain your answer.
(4 + 1) / (1 + 10 + 4 + 4 + 1) = 5 / 20 = 0.25

b. Indicate the interval where the median of this sample can lie. Explain your answer.
20 / 2 = 10
10 + 1 = 11
(-15, -10) -- 1
(-5, 0) -- 10
the interval where the median of this sample lied is (-5, 0).

c. How the histogram will change if we add an element $7$ to the sample? Explain your answer.
the frequency of the interval (5, 10) will extend to 5

### Problem 2

a. Look at the histograms of two samples. They illustrate the distribution of normalized average reaction time to frequent words (in ms) in two groups of people.

```{r, echo=FALSE, fig.height=3}
set.seed(1234)
sample1 <- rnorm(100, 500, 50)
sample2 <- rnorm(100, 500, 20)
hist(sample1, col="lightblue", xlim=c(100, 900))
```

```{r, echo=FALSE, fig.height=3}
hist(sample2, col="lightgreen", xlim=c(100, 900))
```

Which of the samples has a larger variance? Explain your answer.
Sample 1 has a larger variance.
Sample 1 has a larger frequency range than Sample 2: sample 1 from 0 to 40, sample 2 from 0 to 30;
The frequency distribution of sample 1 is more scattered than that of sample 2: frequency of internal 450 - 500 in sample 1 is about 40, but frequency of internal 480 - 520 in sample 2 is about 70

interval 范围比较大
与平均值的差值较大

\newpage

b. Look at the histograms of two samples.

```{r, echo=FALSE, fig.height=8}
set.seed(12)
sample1 <- rnorm(100, 3, 5)
sample2 <- rnorm(100, 10, 0.8)
par(mfrow=c(2,2))
hist(sample1, col="lightblue", xlim=c(-40, 40))
hist(sample2, col="lightgreen", xlim=c(5, 15))
```

Which of the samples has a larger variance? Explain your answer.
sample 1 has a larger variance.
The range of sample 1 is much larger than sample 2: as the total number of each sample is the same, the sum of the three most frequent intervals in sample 1 is 6-0 = 6, the sum of frequencies is about 50, and in sample 2 is about 11 - 9.5 = 1.5 and 60.

## Part 2

*Do not use R (RStudio) to solve problems in Part 2. Answers for problem 3 will be evaluated. Please paste YES or NO into (empty) code blocks and explain you answer below the block.*

### Problem 3

Below is the histogram of the number of mistakes students made while writing an examination essay in English. Look at the histogram and answer the questions.

```{r, echo=FALSE, fig.height=4}
set.seed(2)
mistakes <- rbeta(1000, 2, 7)*100
hist(mistakes, breaks=60, col= "deepskyblue", ylim=c(0, 40))
```

### 3.1
Is it true that $50$% students made more than $35$ mistakes?
```
NO
```
Explain your answer below:
the median of this sample is around 20, so the frequency of students that made more than 35 mistakes is less than 50%.

### 3.2
Is it true that most students made no more than $10$ mistakes?
```
NO
```
Explain your answer below:
The most frequent sample interval is from about 10 to about 20

### 3.3
Which of the following values is closer to be the median of `mistakes`: $10$, $20$, $30$, $40$?
```
20
```
Explain your answer below:
This sample has a normal distribution, so the median is to the right of the mode, which means about 20.

### Problem 4. Exact binomial test

The null hypothesis is that $p=0$ (i.e. no success is possible). In a dataset, there is only one success out of 1 000 000 observations. Will you reject the null hypothesis? 
```
NO
```
Explain your answer below:
p(dataset) = 1 / 1000000 = 0.0000001 this number is very close to 0, so the null hypothesis holds

## Part 3

*Use R (RStudio) to solve problems in Part 3. Your answers will be evaluated. Please paste R code into R code blocks and explain you answer below the block, if needed. *

### Problem 5 

Here is a sample of respondents' age:  

$44$, $50$, $42$, $64$, $66$, $42$, $72$, $56$, $72$, $54$, $46$, $48$, $48$, $52$, $50$, $66$, $84$.

### 5.1
Arrange them in a vector and call it `age`. 
```{r}
age <- c(44, 50, 42, 64, 66, 42, 72, 56, 72, 54, 46, 48, 48, 52, 50, 66, 84)
class(age)
```

### 5.2
Examine the type of `age` variable (numeric, character, etc).
```
numeric
```

### 5.3
Plot the histogram of the vector `age` with $5$ bins. Change its color to any you want. (Use either R basic or ggplot2 style for plotting.)
```{r}
hist(age, col="red", breaks=5)
```


### Problem 6

Here is a series of words:  
*pie, bar, bar, pie, pie, bar, bar, chart*.

### 6.1
Arrange elements above in a vector and call it `words` 
```{r}
words <- c("pie", "bar", "bar", "pie", "pie", "bar", "bar", "chart")
```

### 6.2
Calculate the relative frequences of values in `words` measured in percent.
```{r}
table(words)/sum(table(words))
```

## Problem 7. Position of verbs in verses

The dataset [“The last words in verses”](https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/poetry_last_in_lines.csv) is based on texts written in the 1820s and 1920s (Corpus of Russian Poetry of the Russian National Corpus). Authors collected only one line per author to keep observations as independent from each other as possible. 

Variables:  
- Decade — decade of creation: 1820s, 1920s.  
- RhymedNwords — the number of words in the rhyming position (usually one word, but there are two words in cases such as _вина бы_ 'I would like to get) wine' (which is rhymed with _жабы_ 'toad', see http://russian-poetry.ru/Poem.php?PoemId=18261)).  
- RhymedNsyl — the number of syllables in the rhyming position.  
- UPoS — part of speech of the last word. 
- LineText — a sampled verse.  
- Author — the author of the text.  

Can we decide that in verses written in 1920s, verbs in the rhyming posision are used differently (more often or less often) than expected for verbs in general? 

Let's assume that the probability for verbs to be used in any position ('in general') is 17% (according to [](http://www.ruscorpora.ru/new/corpora-stat.html) ).

### 7.1 State hypothesis

What is your null hypothesis $H_0$ and what is the alternative hypothesis $H_1$?
```
H0: in verses written in 1920s, verbs are used in the rhyming posision more often than expected
H1: in verses written in 1920s, verbs are used in the rhyming posision less often than expected
```

### 7.2

Read the dataset [“The last words in verses”](https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/poetry_last_in_lines.csv). Filter out the relevant observations from 1920s, calculate the number of verbs observed in the sample, and the sample size. 
```{r}
library(readr)
library(dplyr)
library(tidyverse)
verses <- read_delim("https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/poetry_last_in_lines.csv", delim = "\t")
head(verses)
```
```{r}
verses %>%
  filter(Decade == "1920s") ->
  verses_1920s

sum(verses_1920s$RhymedNwords)
```
```{r}
verses_1920s %>% 
  filter(UPoS == "VERB") -> verb_1920s

sum(verb_1920s$RhymedNwords)
```

### 7.3

Use an exact binomial test to calculate p-value. 
```{r}
P_1920verb <- 39 / 283
P_1920verb
```

### 7.4 Interpret results

Give your interpretation of obtained p-value. Answer the initial question: Can we decide that in verses written in 1920s, verbs are used in the rhyming posision more often or less often than expected?
```
in verses written in 1920s, verbs are used in the rhyming posision less often than expected
```

### 7.5 
*(A bonus problem, extra points in evaluation)*.
Repeat 2.3 for verses written in the 1820s.
```{r}
verses %>%
  filter(Decade == "1820s") ->
  verses_1820s

sum(verses_1820s$RhymedNwords)
```
```{r}
verses_1820s %>% 
  filter(UPoS == "VERB") -> verb_1820s

sum(verb_1820s$RhymedNwords)
```
```{r}
P_1820verb <- 15 / 87
P_1820verb
```

Write down your general conlusions about data provided for both 1920s and 1820s data.
```
in verses written in 1920s, verbs are used in the rhyming posision less often than expected, and the deviation from general probability is large;
in verses written in 1920s, verbs are used in the rhyming posision more often than expected, but the deviation from general probability is not large
the probability for verbs to be used in any position ('in general') is 17%
```

### Problem 8. One-sample t-test
Using Icelandic data on vowel duration from seminar [Link](https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/icelandic.csv) test the null hypothesis that the population mean of vowel duration in speaker `shg05` equals 73 (ms).
To perform a one-sample t-test, you can use the following example of R code: 
```
t.test(sample, mu = 7725) # mu is a population mean
```
```{r}
df <- read_csv("https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/icelandic.csv")
```

### 8.1
Write down a two-tailed alternative hypothesis.
```
H1: the population mean of sonorant duration in speaker `tt01` is in the range of 140 - 145 (ms)
```

### 8.2
Perform a one-sample t-test. 
```{r}
t.test(df[df$speaker == 'tt01',]$sonorant.dur, mu=141)
```

### 8.3
Interpret results. 
```
0.015% that the population mean of sonorant duration in speaker `tt01` equals 141 (ms)
```

#### Supplementary reading
Use of exact binomial test in lingiistic research:  

* Gries, Stefan Th. "Phonological similarity in multi-word units." Cognitive Linguistics 22.3 (2011): 491-510. [Link](http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.186.7412&rep=rep1&type=pdf)  
Stefan Gries proves that alliteration is observed in multi-word expressions more often than in general.

* Harald Bayen (2008: 51-52) evaluates the probability of observing exactly one occurrence of the word _hare_ in the corpus sample of 1 mln words given its estimated frequency of 8.23 words per million according to the SELEX frequency database.

On measures of central tendency: 

* Levshina 2015, Chapter 3 (p. 48); Gries 2009, Chapter 1.3 (p. 116). 

On t-test: 

* Gries 2009, Chapter 3 (p. 198). 